---
title: "Madison_GI_Stats"
output: html_document
date: "2025-03-19"
---

To DO:

-Write function in each code block
-In ANCOVA functions, return data frame with site, slope p, slope sig (yes/no), intercept p, intercept sig (yes/no)
-Create table outputs
-Look into 2 way ANOVA
-Run all years in one ANCOVA

Read in data file
```{r}
# change path/file name as necessary 
# skips first row
gi_data <- read.csv("Madison_GI_all.csv", skip = 1)
gi_df <- data.frame(gi_data)
```

Convert values to natural log
```{r}
# Volumes
ln_seg_vol <- log(as.numeric(gi_df$Segoe.AV))
ln_west_vol <- log(as.numeric(gi_df$Westmorland.AV))
ln_glen_vol <- log(as.numeric(gi_df$Glen.AV))
ln_greg_vol <- log(as.numeric(gi_df$Gregory.AV))

# Peak Q
ln_seg_q <- log(as.numeric(gi_df$Segoe.AV.2))
ln_west_q <- log(as.numeric(gi_df$Westmorland.AV.2))
ln_glen_q <- log(as.numeric(gi_df$Glen.AV.2))
ln_greg_q <- log(as.numeric(gi_df$Gregory..AV))
```

Add ln values to data frame
```{r}
gi_df_ln <- cbind(gi_df, ln_seg_vol, ln_west_vol, ln_glen_vol, ln_greg_vol, ln_seg_q, ln_west_q, ln_glen_q, ln_greg_q)
```

Linear Regressions
```{r}
# Glen

# 2020
lm_glen_20 <- lm(ln_glen_vol[1:42] ~ ln_seg_vol[1:42], data = gi_df_ln)
# 2021
lm_glen_21 <- lm(ln_glen_vol[43:84] ~ ln_seg_vol[43:84], data = gi_df_ln)
# 2022
lm_glen_22 <- lm(ln_glen_vol[85:126] ~ ln_seg_vol[85:126], data = gi_df_ln)
# 2023
lm_glen_23 <- lm(ln_glen_vol[127:207] ~ ln_seg_vol[127:207], data = gi_df_ln)

# Plot
# 2020
par(mar = c(4,4,4,1)) # bottom, left, top, right
plot(ln_seg_vol[1:42], ln_glen_vol[1:42], pch=19, col="red", main="Regression of Glen by Segoe", ylab="Glen", xlab="Segoe", ylim=c(0,14), xlim=c(4,13))
# 2021
points(ln_seg_vol[43:84], ln_glen_vol[43:84], pch=19, col="dodgerblue3")
# 2022
points(ln_seg_vol[85:126], ln_glen_vol[85:126], pch=19, col="limegreen")
# 2023
points(ln_seg_vol[127:207], ln_glen_vol[127:207], pch=19, col="gold")
# Regression line 2020
abline(lm_glen_20, lwd=2, col="red")
# Regression line 2021
abline(lm_glen_21, lwd=2, col="dodgerblue3")
# Regression line 2022
abline(lm_glen_22, lwd=2, col="limegreen")
# Regression line 2023
abline(lm_glen_23, lwd=2, col="gold")
grid()
leg.txt <- c("2020", "2021", "2022", "2023")
legend("bottomright", leg.txt, pch=19, col=c("red", "dodgerblue3", "limegreen", "gold"), horiz=TRUE)

# Westmorland

# 2020
lm_west_20 <- lm(ln_west_vol[1:42] ~ ln_seg_vol[1:42], data = gi_df_ln)
# 2021
lm_west_21 <- lm(ln_west_vol[43:84] ~ ln_seg_vol[43:84], data = gi_df_ln)
# 2022
lm_west_22 <- lm(ln_west_vol[85:126] ~ ln_seg_vol[85:126], data = gi_df_ln)
# 2023
lm_west_23 <- lm(ln_west_vol[127:207] ~ ln_seg_vol[127:207], data = gi_df_ln)

# Plot
# 2020
par(mar = c(4,4,4,1)) # bottom, left, top, right
plot(ln_seg_vol[1:42], ln_west_vol[1:42], pch=19, col="red", main="Regression of Westmorland by Segoe", ylab="Westmorland", xlab="Segoe", ylim=c(0,14), xlim=c(4,13))
# 2021
points(ln_seg_vol[43:84], ln_west_vol[43:84], pch=19, col="dodgerblue3")
# 2022
points(ln_seg_vol[85:126], ln_west_vol[85:126], pch=19, col="limegreen")
# 2023
points(ln_seg_vol[127:207], ln_west_vol[127:207], pch=19, col="gold")
# Regression line 2020
abline(lm_west_20, lwd=2, col="red")
# Regression line 2021
abline(lm_west_21, lwd=2, col="dodgerblue3")
# Regression line 2022
abline(lm_west_22, lwd=2, col="limegreen")
# Regression line 2023
abline(lm_west_23, lwd=2, col="gold")
grid()
leg.txt <- c("2020", "2021", "2022", "2023")
legend("bottomright", leg.txt, pch=19, col=c("red", "dodgerblue3", "limegreen", "gold"), horiz=TRUE)

# Gregory

# 2020
lm_greg_20 <- lm(ln_greg_vol[1:42] ~ ln_seg_vol[1:42], data = gi_df_ln)
# 2021
lm_greg_21 <- lm(ln_greg_vol[43:84] ~ ln_seg_vol[43:84], data = gi_df_ln)
# 2022
lm_greg_22 <- lm(ln_greg_vol[85:126] ~ ln_seg_vol[85:126], data = gi_df_ln)
# 2023
lm_greg_23 <- lm(ln_greg_vol[127:207] ~ ln_seg_vol[127:207], data = gi_df_ln)

# Plot
# 2020
par(mar = c(4,4,4,1)) # bottom, left, top, right
plot(ln_seg_vol[1:42], ln_greg_vol[1:42], pch=19, col="red", main="Regression of Gregory by Segoe", ylab="Gregory", xlab="Segoe", ylim=c(0,14), xlim=c(4,13))
# 2021
points(ln_seg_vol[43:84], ln_greg_vol[43:84], pch=19, col="dodgerblue3")
# 2022
points(ln_seg_vol[85:126], ln_greg_vol[85:126], pch=19, col="limegreen")
# 2023
points(ln_seg_vol[127:207], ln_greg_vol[127:207], pch=19, col="gold")
# Regression line 2020
abline(lm_greg_20, lwd=2, col="red")
# Regression line 2021
abline(lm_greg_21, lwd=2, col="dodgerblue3")
# Regression line 2022
abline(lm_greg_22, lwd=2, col="limegreen")
# Regression line 2023
abline(lm_greg_23, lwd=2, col="gold")
grid()
leg.txt <- c("2020", "2021", "2022", "2023")
legend("bottomright", leg.txt, pch=19, col=c("red", "dodgerblue3", "limegreen", "gold"), horiz=TRUE)

```

Get % Differences
```{r}
# Glen

# 2020/2021
glen_20_21_pred_ln <- lm_glen_20$coefficients[2]*gi_df_ln$ln_seg_vol[43:84] + lm_glen_20$coefficients[1]
glen_20_21_pred <- exp(glen_20_21_pred_ln)

glen_diff_20_21 <- (mean(gi_df_ln$Glen.AV[43:84]) - mean(glen_20_21_pred))/mean(glen_20_21_pred)

# 2020/2022
glen_20_22_pred_ln <- lm_glen_20$coefficients[2]*gi_df_ln$ln_seg_vol[85:126] + lm_glen_20$coefficients[1]
glen_20_22_pred <- exp(glen_20_22_pred_ln)

glen_diff_20_22 <- (mean(gi_df_ln$Glen.AV[85:126]) - mean(glen_20_22_pred))/mean(glen_20_22_pred)

# 2020/2023
glen_20_23_pred_ln <- lm_glen_20$coefficients[2]*gi_df_ln$ln_seg_vol[127:207] + lm_glen_20$coefficients[1]
glen_20_23_pred <- exp(glen_20_23_pred_ln)

glen_diff_20_23 <- (mean(gi_df_ln$Glen.AV[127:207]) - mean(glen_20_23_pred))/mean(glen_20_23_pred)

# 2021/2022
glen_21_22_pred_ln <- lm_glen_21$coefficients[2]*gi_df_ln$ln_seg_vol[85:126] + lm_glen_21$coefficients[1]
glen_21_22_pred <- exp(glen_21_22_pred_ln)

glen_diff_21_22 <- (mean(gi_df_ln$Glen.AV[85:126]) - mean(glen_21_22_pred))/mean(glen_21_22_pred)

# 2021/2023
glen_21_23_pred_ln <- lm_glen_21$coefficients[2]*gi_df_ln$ln_seg_vol[127:207] + lm_glen_21$coefficients[1]
glen_21_23_pred <- exp(glen_21_23_pred_ln)

glen_diff_21_23 <- (mean(gi_df_ln$Glen.AV[127:207]) - mean(glen_21_23_pred))/mean(glen_21_23_pred)

# 2022/2023
glen_22_23_pred_ln <- lm_glen_22$coefficients[2]*gi_df_ln$ln_seg_vol[127:207] + lm_glen_22$coefficients[1]
glen_22_23_pred <- exp(glen_22_23_pred_ln)

glen_diff_22_23 <- (mean(gi_df_ln$Glen.AV[127:207]) - mean(glen_22_23_pred))/mean(glen_22_23_pred)

# Westmorland

# 2020/2021
west_20_21_pred_ln <- lm_west_20$coefficients[2]*gi_df_ln$ln_seg_vol[43:84] + lm_west_20$coefficients[1]
west_20_21_pred <- exp(west_20_21_pred_ln)

west_diff_20_21 <- (mean(gi_df_ln$Westmorland.AV[43:84]) - mean(west_20_21_pred))/mean(west_20_21_pred)

# 2020/2022
west_20_22_pred_ln <- lm_west_20$coefficients[2]*gi_df_ln$ln_seg_vol[85:126] + lm_west_20$coefficients[1]
west_20_22_pred <- exp(west_20_22_pred_ln)

west_diff_20_22 <- (mean(gi_df_ln$Westmorland.AV[85:126]) - mean(west_20_22_pred))/mean(west_20_22_pred)

# 2020/2023
west_20_23_pred_ln <- lm_west_20$coefficients[2]*gi_df_ln$ln_seg_vol[127:207] + lm_west_20$coefficients[1]
west_20_23_pred <- exp(west_20_23_pred_ln)

west_diff_20_23 <- (mean(gi_df_ln$Westmorland.AV[127:207]) - mean(west_20_23_pred))/mean(west_20_23_pred)

# 2021/2022
west_21_22_pred_ln <- lm_west_21$coefficients[2]*gi_df_ln$ln_seg_vol[85:126] + lm_west_21$coefficients[1]
west_21_22_pred <- exp(west_21_22_pred_ln)

west_diff_21_22 <- (mean(gi_df_ln$Westmorland.AV[85:126]) - mean(west_21_22_pred))/mean(west_21_22_pred)

# 2021/2023
west_21_23_pred_ln <- lm_west_21$coefficients[2]*gi_df_ln$ln_seg_vol[127:207] + lm_west_21$coefficients[1]
west_21_23_pred <- exp(west_21_23_pred_ln)

west_diff_21_23 <- (mean(gi_df_ln$Westmorland.AV[127:207]) - mean(west_21_23_pred))/mean(west_21_23_pred)

# 2022/2023
west_22_23_pred_ln <- lm_west_22$coefficients[2]*gi_df_ln$ln_seg_vol[127:207] + lm_west_22$coefficients[1]
west_22_23_pred <- exp(west_22_23_pred_ln)

west_diff_22_23 <- (mean(gi_df_ln$Westmorland.AV[127:207]) - mean(west_22_23_pred))/mean(west_22_23_pred)

# Gregory

# 2020/2021
greg_20_21_pred_ln <- lm_greg_20$coefficients[2]*gi_df_ln$ln_seg_vol[43:84] + lm_greg_20$coefficients[1]
greg_20_21_pred <- exp(greg_20_21_pred_ln)

greg_diff_20_21 <- (mean(gi_df_ln$Gregory.AV[43:84]) - mean(greg_20_21_pred))/mean(greg_20_21_pred)

# 2020/2022
greg_20_22_pred_ln <- lm_greg_20$coefficients[2]*gi_df_ln$ln_seg_vol[85:126] + lm_greg_20$coefficients[1]
greg_20_22_pred <- exp(greg_20_22_pred_ln)

greg_diff_20_22 <- (mean(gi_df_ln$Gregory.AV[85:126]) - mean(greg_20_22_pred))/mean(greg_20_22_pred)

# 2020/2023
greg_20_23_pred_ln <- lm_greg_20$coefficients[2]*gi_df_ln$ln_seg_vol[127:207] + lm_greg_20$coefficients[1]
greg_20_23_pred <- exp(greg_20_23_pred_ln)

greg_diff_20_23 <- (mean(gi_df_ln$Gregory.AV[127:207]) - mean(greg_20_23_pred))/mean(greg_20_23_pred)

# 2021/2022
greg_21_22_pred_ln <- lm_greg_21$coefficients[2]*gi_df_ln$ln_seg_vol[85:126] + lm_greg_21$coefficients[1]
greg_21_22_pred <- exp(greg_21_22_pred_ln)

greg_diff_21_22 <- (mean(gi_df_ln$Gregory.AV[85:126]) - mean(greg_21_22_pred))/mean(greg_21_22_pred)

# 2021/2023
greg_21_23_pred_ln <- lm_greg_21$coefficients[2]*gi_df_ln$ln_seg_vol[127:207] + lm_greg_21$coefficients[1]
greg_21_23_pred <- exp(greg_21_23_pred_ln)

greg_diff_21_23 <- (mean(gi_df_ln$Gregory.AV[127:207]) - mean(greg_21_23_pred))/mean(greg_21_23_pred)

# 2022/2023
greg_22_23_pred_ln <- lm_greg_22$coefficients[2]*gi_df_ln$ln_seg_vol[127:207] + lm_greg_22$coefficients[1]
greg_22_23_pred <- exp(greg_22_23_pred_ln)

greg_diff_22_23 <- (mean(gi_df_ln$Gregory.AV[127:207]) - mean(greg_22_23_pred))/mean(greg_22_23_pred)
```

ANCOVA Glen
```{r}
# 2020/2021
ancova_glen_20_21 <- aov(ln_glen_vol[1:84] ~ as.factor(Year[1:84]) * ln_seg_vol[1:84] , data = gi_df_ln)
Anova(ancova_glen_20_21, type="III")

# anova_glen_20_21 <- aov(ln_glen_vol[1:84] ~ as.factor(Year[1:84]), data = gi_df_ln)
# summary(anova_glen_20_21)

TukeyHSD(ancova_glen_20_21, conf.level=0.9)

# 2020/2022
ancova_glen_20_22 <- aov(ln_glen_vol[c(1:42, 85:126)] ~ as.factor(Year[c(1:42, 85:126)]) * ln_seg_vol[c(1:42, 85:126)] , data = gi_df_ln)
Anova(ancova_glen_20_22, type="II")

TukeyHSD(ancova_glen_20_22, conf.level=0.9)

# 2020/2023
ancova_glen_20_23 <- aov(ln_glen_vol[c(1:42, 127:207)] ~ as.factor(Year[c(1:42, 127:207)]) * ln_seg_vol[c(1:42, 127:207)] , data = gi_df_ln)
Anova(ancova_glen_20_23, type="II")

TukeyHSD(ancova_glen_20_23, "Year", conf.level=0.9)

# 2021/2022
ancova_glen_21_22 <- aov(ln_glen_vol[43:126] ~ as.factor(Year[43:126]) * ln_seg_vol[43:126] , data = gi_df_ln)
Anova(ancova_glen_21_22, type="II")

TukeyHSD(ancova_glen_21_22, conf.level=0.9)

# 2021/2023
ancova_glen_21_23 <- aov(ln_glen_vol[c(43:84, 127:207)] ~ as.factor(Year[c(43:84, 127:207)]) * ln_seg_vol[c(43:84, 127:207)] , data = gi_df_ln)
Anova(ancova_glen_21_23, type="II")

TukeyHSD(ancova_glen_21_23, "Year", conf.level=0.9)

# 2022/2023
ancova_glen_22_23 <- aov(ln_glen_vol[85:207] ~ as.factor(Year[85:207]) * ln_seg_vol[85:207] , data = gi_df_ln)
Anova(ancova_glen_22_23, type="II")

TukeyHSD(ancova_glen_22_23, "Year", conf.level=0.9)

```

ANCOVA Westmorland
```{r}
# 2020/2021
ancova_west_20_21 <- aov(ln_west_vol[1:84] ~ as.factor(Year[1:84]) * ln_seg_vol[1:84] , data = gi_df_ln)
Anova(ancova_west_20_21, type="III")

# anova_west_20_21 <- aov(ln_west_vol[1:84] ~ as.factor(Year[1:84]), data = gi_df_ln)
# summary(anova_west_20_21)

TukeyHSD(ancova_west_20_21, conf.level=0.9)

# 2020/2022
ancova_west_20_22 <- aov(ln_west_vol[c(1:42, 85:126)] ~ as.factor(Year[c(1:42, 85:126)]) * ln_seg_vol[c(1:42, 85:126)] , data = gi_df_ln)
Anova(ancova_west_20_22, type="II")

TukeyHSD(ancova_west_20_22, conf.level=0.9)

# 2020/2023
ancova_west_20_23 <- aov(ln_west_vol[c(1:42, 127:207)] ~ as.factor(Year[c(1:42, 127:207)]) * ln_seg_vol[c(1:42, 127:207)] , data = gi_df_ln)
Anova(ancova_west_20_23, type="II")

TukeyHSD(ancova_west_20_23, "Year", conf.level=0.9)

# 2021/2022
ancova_west_21_22 <- aov(ln_west_vol[43:126] ~ as.factor(Year[43:126]) * ln_seg_vol[43:126] , data = gi_df_ln)
Anova(ancova_west_21_22, type="II")

TukeyHSD(ancova_west_21_22, conf.level=0.9)

# 2021/2023
ancova_west_21_23 <- aov(ln_west_vol[c(43:84, 127:207)] ~ as.factor(Year[c(43:84, 127:207)]) * ln_seg_vol[c(43:84, 127:207)] , data = gi_df_ln)
Anova(ancova_west_21_23, type="II")

TukeyHSD(ancova_west_21_23, "Year", conf.level=0.9)

# 2022/2023
ancova_west_22_23 <- aov(ln_west_vol[85:207] ~ as.factor(Year[85:207]) * ln_seg_vol[85:207] , data = gi_df_ln)
Anova(ancova_west_22_23, type="II")

TukeyHSD(ancova_west_22_23, "Year", conf.level=0.9)

```

ANCOVA Gregory
```{r}
# 2020/2021
ancova_greg_20_21 <- aov(ln_greg_vol[1:84] ~ as.factor(Year[1:84]) * ln_seg_vol[1:84] , data = gi_df_ln)
Anova(ancova_greg_20_21, type="III")

# anova_greg_20_21 <- aov(ln_greg_vol[1:84] ~ as.factor(Year[1:84]), data = gi_df_ln)
# summary(anova_greg_20_21)

TukeyHSD(ancova_greg_20_21, conf.level=0.9)

# 2020/2022
ancova_greg_20_22 <- aov(ln_greg_vol[c(1:42, 85:126)] ~ as.factor(Year[c(1:42, 85:126)]) * ln_seg_vol[c(1:42, 85:126)] , data = gi_df_ln)
Anova(ancova_greg_20_22, type="II")

TukeyHSD(ancova_greg_20_22, conf.level=0.9)

# 2020/2023
ancova_greg_20_23 <- aov(ln_greg_vol[c(1:42, 127:207)] ~ as.factor(Year[c(1:42, 127:207)]) * ln_seg_vol[c(1:42, 127:207)] , data = gi_df_ln)
Anova(ancova_greg_20_23, type="II")

TukeyHSD(ancova_greg_20_23, "Year", conf.level=0.9)

# 2021/2022
ancova_greg_21_22 <- aov(ln_greg_vol[43:126] ~ as.factor(Year[43:126]) * ln_seg_vol[43:126] , data = gi_df_ln)
Anova(ancova_greg_21_22, type="II")

TukeyHSD(ancova_greg_21_22, conf.level=0.9)

# 2021/2023
ancova_greg_21_23 <- aov(ln_greg_vol[c(43:84, 127:207)] ~ as.factor(Year[c(43:84, 127:207)]) * ln_seg_vol[c(43:84, 127:207)] , data = gi_df_ln)
Anova(ancova_greg_21_23, type="II")

TukeyHSD(ancova_greg_21_23, "Year", conf.level=0.9)

# 2022/2023
ancova_greg_22_23 <- aov(ln_greg_vol[85:207] ~ as.factor(Year[85:207]) * ln_seg_vol[85:207] , data = gi_df_ln)
Anova(ancova_greg_22_23, type="II")

TukeyHSD(ancova_greg_22_23, "Year", conf.level=0.9)

```

Create data frame (table) of results
```{r}
# Columns needed: Test Site, Year Comparison, Slope p value, Slope significant?, Intercept p value, Intercept significant, % Change

year_comps <- c("2020 vs 2021", "2020 vs 2022", "2020 vs 2023", "2021 vs 2022", "2021 vs 2023", "2022 vs 2023")
```

