---
title: "Nonlinear_Modeling"
output: html_document
date: "2024-07-22"
---

```{r}
library(mgcv)
library(splines)
library(caret)
library(gamclass)
library(ggplot2)
library(car)
library(randomForest)
library(mice)
#library(gam)
#detach("package:gam", unload = TRUE)

df_all <- read.csv("linear model data_V2_all_impute.csv")
df_pred_removed <- read.csv("linear model data_V2_predictors_removed_pca.csv")
df_res_removed <- read.csv("linear model data_V2_res_removed_pca_rf_binom.csv")
df_subset <- read.csv("lm_data_all_subset_impute.csv")
```

Impute missing data using MICE:
```{r}
imputed_data <- mice(df_subset, m = 5, method = "pmm", seed = 123)

#summary(imputed_data)

imputed_df <- complete(imputed_data, 1)
```
Add imputed data to dataframe
```{r}
# Function to replace NA values in the original dataframe with imputed values
replace_na <- function(original, imputed) {
  original[is.na(original)] <- imputed[is.na(original)]
  return(original)
}

# Apply the function to each column with missing values
for (col in names(df_subset)) {
  if (any(is.na(df_subset[[col]]))) {
    df_subset[[col]] <- replace_na(df_subset[[col]], imputed_df[[col]])
  }
}
```

Scale predictor vars:
```{r}
predictors <- df_res_removed[, -1]

scaled_predictors <- scale(predictors)

df_res_removed_scaled <- data.frame(response = df_res_removed$Average.infiltration.rate, scaled_predictors)
```

Outliers
```{r}
inf_outliers <- abs(scale(df_res_removed$Average.infiltration.rate)) > 3
df_inf_outliers <- df_res_removed[!inf_outliers, ]
```

Correlation Maxtrix
```{r}
corr_matrix <- cor(df_res_removed[3:43])

write.csv(corr_matrix, file = "correlation.matrix.csv")
```

Check VIF
```{r}
y <- df_subset$Average.infiltration.rate

# Area+Contributing.area+Drainage.Area.Ratio+Age+Ponding.depth+Inlet.condition+Outlet.condition+Underdrain+X..Residential+X..Commercial+X..Industrial+X..Institutional+X..Vegetation_shrub+X..Vegetation_tree+X..Vegetation_grass+X..Vegetation_prairie+X..Vegetation_bare+X..Vegetation_other+Underlying.soil_..sand+Underlying.soil_..silt+Underlying.soil_..clay+Underlying.soil_..urban.disturbed+Underlying.soil_..other.undefined+Underlying.soil_..organic.matter+Evidence.of.poor.infiltration+X..of.surface.with.poor.infiltration+Vegetation.condition+Inlet.feature_Curb.Cut+Inlet.feature_Other+Inlet.feature_Overland.Flow+Inlet.feature_Pipe+Outlet.feature_Other+Outlet.feature_Overland.Flow+Outlet.feature_Pipe+Outlet.feature_Riser.Pipe+Engineered.soil.depth+Engineered.soil._..sand+Engineered.soil._..silt+Engineered.soil._..clay+Engineered.soil._..compost+Engineered.soil._..other

# Pared down: Drainage.Area.Ratio+Age+Ponding.depth+Underdrain+X..Residential+X..Vegetation_shrub+X..Vegetation_tree+X..Vegetation_grass+X..Vegetation_other+X..of.surface.with.poor.infiltration+Vegetation.condition+Engineered.soil.depth+Engineered.soil._..sand+Engineered.soil._..compost+Soil_PC_1+Soil_PC_2 

# Deleted: X..Industrial, X..Vegetation_bare, Underlying.soil_..other.undefined, Inlet.feature_Other, Outlet.feature_Other

vif_model <- lm(Average.infiltration.rate ~ Drainage.Area.Ratio+Age+Ponding.depth+Underdrain+X..Residential+X..of.surface.with.poor.infiltration+Vegetation.condition+Engineered.soil.depth+Engineered.soil._..sand+Engineered.soil._..compost+Soil_PC_1+Soil_PC_2, data = df_res_removed)

#create vector of VIF values
vif_values <- vif(vif_model)

par(mar=c(4,20,4,4))
#create horizontal bar chart to display each VIF value
barplot(vif_values, horiz = TRUE, cex.names = 2, cex.axis = 2, las=1, col = "steelblue")

#add vertical lines at 5 and 10
abline(v = 5, lwd = 2, lty = 2)
abline(v = 10, lwd = 2, lty = 2)
```

Random forest for variable selection
```{r}
#make this example reproducible
set.seed(1)

#fit the random forest model
model <- randomForest(
  formula = Average.infiltration.rate ~ .,
  data = df_subset
)

#display fitted model
model

#find number of trees that produce lowest test MSE
which.min(model$mse)

#find RMSE of best model
sqrt(model$mse[which.min(model$mse)])

# plot the test MSE by number of trees
plot(model)

# produce variable importance plot
varImpPlot(model)

model_tuned <- tuneRF( x=df_subset[,-1], y=df_subset$Average.infiltration.rate, ntreeTry=500, mtryStart=4, stepFactor=1.5, improve=0.01, trace=FALSE)

#display fitted model
model_tuned


```

```{r}
model <- gam(Average.infiltration.rate ~ Drainage.Area.Ratio + s(Age) + Underdrain +
    s(X..Vegetation_shrub)    + Ponding.depth + s(Engineered.soil.depth)  + s(Engineered.soil._..sand) + Soil_PC_1 +
    s(Soil_PC_2), select=TRUE, data = df_subset)

summary(model)

#gam.check(model)
```

Step GAM for variable selection
```{r}
full_model <- gam(Average.infiltration.rate ~ s(Drainage.Area.Ratio) + s(Age) + s(Ponding.depth)  + s(X..Residential) + s(X..Vegetation_shrub) + s(X..Vegetation_tree) + X..Vegetation_grass + s(X..Vegetation_prairie) + s(X..Vegetation_other) + s(Soil_PC_1) + s(Soil_PC_2)  + X..of.surface.with.poor.infiltration   +  s(Engineered.soil.depth) + s(Engineered.soil._..sand) + Engineered.soil._..compost, data=df_res_removed)

#summary(full_model)

scope_list <- gam.scope(df_res_removed, response=1, smoother="s")

step.Gam(full_model, scope_list)
#scope_list <- list("Drainage.Area.Ratio" = ~1 + Drainage.Area.Ratio + s(Drainage.Area.Ratio, df=2), s(Drainage.Area.Ratio, df=3), s(Drainage.Area.Ratio, df=4), s(Drainage.Area.Ratio, df=5), "Age" = ~1 + Age + s(Age, df=2), s(Age, df=3), s(Age, df=4), s(Age, df=5), "Ponding.depth" = ~1 + Ponding.depth + s(Ponding.depth, df=2), s(Ponding.depth, df=3), s(Ponding.depth, df=4), s(Ponding.depth, df=5), "Underdrain" = ~1 + Underdrain + s(Underdrain, df=2), s(Underdrain, df=3), s(Underdrain, df=4), s(Underdrain, df=5), "X..Residential" = ~1 + X..Residential + s(X..Residential, df=2), s(X..Residential, df=3), s(X..Residential, df=4), s(X..Residential, df=5), "X..Vegetation_shrub" = ~1 + X..Vegetation_shrub + s(X..Vegetation_shrub, df=2), s(X..Vegetation_shrub, df=3), s(X..Vegetation_shrub, df=4), s(X..Vegetation_shrub, df=5), "X..Vegetation_tree" = ~1 + X..Vegetation_tree + s(X..Vegetation_tree, df=2), s(X..Vegetation_tree, df=3), s(X..Vegetation_tree, df=4), s(X..Vegetation_tree, df=5), "X..Vegetation_grass" = ~1 + X..Vegetation_grass + s(X..Vegetation_grass, df=2), s(X..Vegetation_grass, df=3), s(X..Vegetation_grass, df=4), s(X..Vegetation_grass, df=5), "X..Vegetation_prairie" = ~1 + X..Vegetation_prairie + s(X..Vegetation_prairie, df=2), s(X..Vegetation_prairie, df=3), s(X..Vegetation_prairie, df=4), s(X..Vegetation_prairie, df=5), "X..Vegetation_other" = ~1 + X..Vegetation_other + s(X..Vegetation_other, df=2), s(X..Vegetation_other, df=3), s(X..Vegetation_other, df=4), s(X..Vegetation_other, df=5), "Soil_PC_1" = ~1 + Soil_PC_1 + s(Soil_PC_1, df=2), s(Soil_PC_1, df=3), s(Soil_PC_1, df=4), s(Soil_PC_1, df=5), "Soil_PC_2" = ~1 + Soil_PC_2 + s(Soil_PC_2, df=2), s(Soil_PC_2, df=3), s(Soil_PC_2, df=4), s(Soil_PC_2, df=5), "X..of.surface.with.poor.infiltration" = ~1 + X..of.surface.with.poor.infiltration + s(X..of.surface.with.poor.infiltration, df=2), s(X..of.surface.with.poor.infiltration, df=3), s(X..of.surface.with.poor.infiltration, df=4), s(X..of.surface.with.poor.infiltration, df=5), "Vegetation.condition" = ~1 + Vegetation.condition + s(Vegetation.condition, df=2), s(Vegetation.condition, df=3), s(Vegetation.condition, df=4), s(Vegetation.condition, df=5), "Engineered.soil.depth" = ~1 + Engineered.soil.depth + s(Engineered.soil.depth, df=2), s(Engineered.soil.depth, df=3), s(Engineered.soil.depth, df=4), s(Engineered.soil.depth, df=5), "Engineered.soil._..sand" = ~1 + Engineered.soil._..sand + s(Engineered.soil._..sand, df=2), s(Engineered.soil._..sand, df=3), s(Engineered.soil._..sand, df=4), s(Engineered.soil._..sand, df=5), "Engineered.soil._..compost" = ~1 + Engineered.soil._..compost + s(Engineered.soil._..compost, df=2), s(Engineered.soil._..compost, df=3), s(Engineered.soil._..compost, df=4), s(Engineered.soil._..compost, df=5))
```

GAM
```{r}
# All smoothed
#s(Area) + s(Contributing.area) + s(Drainage.Area.Ratio) + s(Age) + s(Ponding.depth) + s(Inlet.condition) + s(Outlet.condition) + s(Underdrain) + s(X..Residential) + s(X..Commercial) + s(X..Industrial) + s(X..Institutional) + s(X..Vegetation_shrub) + s(X..Vegetation_tree) + s(X..Vegetation_grass) + s(X..Vegetation_prairie) + s(X..Vegetation_bare) + s(X..Vegetation_other) + s(Underlying.soil_..sand) + s(Underlying.soil_..silt) + s(Underlying.soil_..clay) + s(Underlying.soil_..urban.disturbed) + s(Underlying.soil_..other.undefined) + s(Underlying.soil_..organic.matter) + s(Evidence.of.poor.infiltration) + s(X..of.surface.with.poor.infiltration) + s(Vegetation.condition) + s(Inlet.feature_Curb.Cut) + s(Inlet.feature_Other) + s(Inlet.feature_Overland.Flow) + s(Inlet.feature_Pipe) + s(Outlet.feature_Other) + s(Outlet.feature_Overland.Flow) + s(Outlet.feature_Pipe) + s(Outlet.feature_Riser.Pipe) + s(Engineered.soil.depth) + s(Engineered.soil._..sand) + s(Engineered.soil._..silt) + s(Engineered.soil._..clay) + s(Engineered.soil._..compost) + s(Engineered.soil._..other)

#Average.infiltration.rate ~ s(X..Industrial)s(X..Vegetation_bare) + s(X..of.surface.with.poor.infiltration) Inlet.feature_Pipe Inlet.feature_Other s(Engineered.soil._..clay)
#s(Area) + s(Contributing.area) + s(Drainage.Area.Ratio)+ s(Age) + s(Ponding.depth)+ Inlet.condition + Outlet.condition + Underdrain + s(X..Residential) + s(X..Commercial) + s(X..Institutional) + s(X..Vegetation_shrub) + s(X..Vegetation_tree) + s(X..Vegetation_grass) + s(X..Vegetation_prairie) + s(X..Vegetation_other) +  s(Underlying.soil_..sand) + s(Underlying.soil_..silt) + s(Underlying.soil_..clay) + s(Underlying.soil_..urban.disturbed) + s(Underlying.soil_..other.undefined) + s(Underlying.soil_..organic.matter) + Evidence.of.poor.infiltration + Vegetation.condition + Inlet.feature_Curb.Cut + Inlet.feature_Overland.Flow + Inlet.feature_Pipe + Inlet.feature_Other + Outlet.feature_Overland.Flow + Outlet.feature_Pipe + Outlet.feature_Riser.Pipe + Outlet.feature_Other + s(Engineered.soil.depth) + s(Engineered.soil._..sand) + s(Engineered.soil._..silt) + s(Engineered.soil._..clay) + s(Engineered.soil._..compost) + s(Engineered.soil._..other)

#y <- df_inf_outliers$Average.infiltration.rate
# Only used predictor vars that didn't give error: A term has fewer unique covariate combinations than specified maximum degrees of freedom
gam_model <- gam(Average.infiltration.rate ~   s(Drainage.Area.Ratio)  + s(Ponding.depth) + Underdrain + s(X..Residential)  + s(X..Vegetation_tree) + X..Vegetation_grass + s(X..Vegetation_prairie) + s(Soil_PC_1) + s(Soil_PC_2)  + X..of.surface.with.poor.infiltration + Vegetation.condition  +  s(Engineered.soil.depth) + s(Engineered.soil._..sand) + Engineered.soil._..compost
, data = df_subset)

gam.check(gam_model)

#concurvity(gam_model, full = TRUE)

summary(gam_model)

plot(gam_model, all.terms = TRUE, residuals=TRUE, pch=1, cex=1, shade=TRUE, shade.col="lightblue")
```

Plot predictor vs response vars
```{r}
#s(Area) + s(Contributing.area) + s(Drainage.Area.Ratio) + s(Age) + s(Ponding.depth) + s(Inlet.condition) + s(Outlet.condition) + s(Underdrain) + s(X..Residential) + s(X..Commercial) + s(X..Industrial) + s(X..Institutional) + s(X..Vegetation_shrub) + s(X..Vegetation_tree) + s(X..Vegetation_grass) + s(X..Vegetation_prairie) + s(X..Vegetation_bare) + s(X..Vegetation_other) + s(Underlying.soil_..sand) + s(Underlying.soil_..silt) + s(Underlying.soil_..clay) + s(Underlying.soil_..urban.disturbed) + s(Underlying.soil_..other.undefined) + s(Underlying.soil_..organic.matter) + s(Evidence.of.poor.infiltration) + s(X..of.surface.with.poor.infiltration) + s(Vegetation.condition) + s(Inlet.feature_Curb.Cut) + s(Inlet.feature_Other) + s(Inlet.feature_Overland.Flow) + s(Inlet.feature_Pipe) + s(Outlet.feature_Other) + s(Outlet.feature_Overland.Flow) + s(Outlet.feature_Pipe) + s(Outlet.feature_Riser.Pipe) + s(Engineered.soil.depth) + s(Engineered.soil._..sand) + s(Engineered.soil._..silt) + s(Engineered.soil._..clay) + s(Engineered.soil._..compost) + s(Engineered.soil._..other)

plot(df_res_removed$Engineered.soil._..other, df_res_removed$Average.infiltration.rate, pch=16, col="lightblue")
```

GAM with REML Method
```{r}
#s(Area) + s(Contributing.area) + s(Drainage.Area.Ratio) + s(Age) + s(Ponding.depth) + s(Inlet.condition) + s(Outlet.condition) + s(Underdrain) + s(X..Residential) + s(X..Commercial) + s(X..Industrial) + s(X..Institutional) + s(X..Vegetation_shrub) + s(X..Vegetation_tree) + s(X..Vegetation_grass) + s(X..Vegetation_prairie) + s(X..Vegetation_bare) + s(X..Vegetation_other) + s(Underlying.soil_..sand) + s(Underlying.soil_..silt) + s(Underlying.soil_..clay) + s(Underlying.soil_..urban.disturbed) + s(Underlying.soil_..other.undefined) + s(Underlying.soil_..organic.matter) + s(Evidence.of.poor.infiltration) + s(X..of.surface.with.poor.infiltration) + s(Vegetation.condition) + s(Inlet.feature_Curb.Cut) + s(Inlet.feature_Other) + s(Inlet.feature_Overland.Flow) + s(Inlet.feature_Pipe) + s(Outlet.feature_Other) + s(Outlet.feature_Overland.Flow) + s(Outlet.feature_Pipe) + s(Outlet.feature_Riser.Pipe) + s(Engineered.soil.depth) + s(Engineered.soil._..sand) + s(Engineered.soil._..silt) + s(Engineered.soil._..clay) + s(Engineered.soil._..compost) + s(Engineered.soil._..other)

#s(Drainage.Area.Ratio) + s(Age, bs="re") + s(Ponding.depth) + Underdrain + s(X..Residential) + s(X..Vegetation_shrub) + s(X..Vegetation_tree) + s(X..Vegetation_grass) + s(X..Vegetation_prairie) + s(X..Vegetation_other) + s(Soil_PC_1) + s(Soil_PC_2)  + X..of.surface.with.poor.infiltration + Vegetation.condition  +  s(Engineered.soil.depth) + s(Engineered.soil._..sand) + s(Engineered.soil._..compost)

df_res_removed$High.Low.Inf <- as.factor(df_res_removed$High.Low.Inf)

gam_model <- gam(High.Low.Inf ~  Ponding.depth + Underdrain  + X..Vegetation_shrub  + s(X..Vegetation_grass)    + s(Soil_PC_2)     +  s(Engineered.soil.depth), family = binomial, data = df_res_removed)

gam.check(gam_model)

#concurvity(gam_model, full = TRUE)

summary(gam_model)

plot(gam_model, all.terms = TRUE, residuals=TRUE, pch=1, cex=1, shade=TRUE, shade.col="lightblue")


```

Create custom model in caret
```{r}
# Define response variable
y <- df_res_removed$Average.infiltration.rate

# Custom train function for GAM
gam_model <- list(
  type = "Regression",
  library = "mgcv",
  loop = NULL,
  parameters = data.frame(parameter = "parameter", class = "numeric", label = "parameter"),
  grid = function(x, y, len = NULL, search = "grid") data.frame(parameter = 1),
  fit = function(x, y, wts, param, lev, last, classProbs, ...) {
    data <- cbind(x, y = y)
    gam(y ~ s(Area) + s(Contributing.area) + Drainage.Area.Ratio+ s(Age) + s(Ponding.depth)+ Inlet.condition + Outlet.condition + Underdrain + X..Residential + s(X..Commercial) + X..Industrial + X..Institutional + s(X..Vegetation_shrub) + X..Vegetation_tree + X..Vegetation_grass + s(X..Vegetation_prairie)+ X..Vegetation_bare + s(X..Vegetation_other) +  Underlying.soil_..sand + Underlying.soil_..silt + Underlying.soil_..clay + Underlying.soil_..urban.disturbed + Underlying.soil_..other.undefined + s(Underlying.soil_..organic.matter) + Evidence.of.poor.infiltration + X..of.surface.with.poor.infiltration + Vegetation.condition + Inlet.feature_Curb.Cut + Inlet.feature_Overland.Flow + Inlet.feature_Pipe + Inlet.feature_Other + Outlet.feature_Overland.Flow + Outlet.feature_Pipe + Outlet.feature_Riser.Pipe + Outlet.feature_Other + Engineered.soil.depth + s(Engineered.soil._..sand) + Engineered.soil._..silt + Engineered.soil._..clay + Engineered.soil._..compost + s(Engineered.soil._..other), data = df_res_removed, ...)
  },
  predict = function(modelFit, newdata, submodels = NULL) {
    predict(modelFit, newdata)
  },
  prob = NULL
)

# Set control for cross-validation
train_control <- trainControl(method = "cv", number = 5)

# Fit the custom GAM model using caret
set.seed(123)
custom_gam <-train(
  x = df_res_removed[, c("Area", "Contributing.area", "Drainage.Area.Ratio", "Age", "Ponding.depth", "Inlet.condition", "Outlet.condition", "Underdrain", "X..Residential", "X..Commercial", "X..Industrial", "X..Institutional", "X..Vegetation_shrub", "X..Vegetation_tree", "X..Vegetation_grass", "X..Vegetation_prairie", "X..Vegetation_bare", "X..Vegetation_other", "Underlying.soil_..sand", "Underlying.soil_..silt", "Underlying.soil_..clay", "Underlying.soil_..urban.disturbed", "Underlying.soil_..other.undefined", "Underlying.soil_..organic.matter", "Evidence.of.poor.infiltration", "X..of.surface.with.poor.infiltration", "Vegetation.condition", "Inlet.feature_Curb.Cut", "Inlet.feature_Overland.Flow", "Inlet.feature_Pipe", "Inlet.feature_Other", "Outlet.feature_Overland.Flow", "Outlet.feature_Pipe", "Outlet.feature_Riser.Pipe", "Outlet.feature_Other", "Engineered.soil.depth", "Engineered.soil._..sand", "Engineered.soil._..silt", "Engineered.soil._..clay", "Engineered.soil._..compost", "Engineered.soil._..other")],
  y = df_res_removed$Average.infiltration.rate,,
  method = gam_model,
  trControl = train_control
)

# Print results
print(custom_gam)
```

Manual Cross Validation
```{r}
# Define response variable
y <- df_res_removed$Average.infiltration.rate

# Define the GAM model formula
formula <- y ~ s(Area) + s(Contributing.area) + Drainage.Area.Ratio + s(Age) + s(Ponding.depth)+ Inlet.condition + Outlet.condition + Underdrain + X..Residential + s(X..Commercial) + X..Industrial + X..Institutional + s(X..Vegetation_shrub) + X..Vegetation_tree + X..Vegetation_grass + s(X..Vegetation_prairie)+ X..Vegetation_bare + s(X..Vegetation_other) +  Underlying.soil_..sand + Underlying.soil_..silt + Underlying.soil_..clay + Underlying.soil_..urban.disturbed + Underlying.soil_..other.undefined + s(Underlying.soil_..organic.matter) + Evidence.of.poor.infiltration + X..of.surface.with.poor.infiltration + Vegetation.condition + Inlet.feature_Curb.Cut + Inlet.feature_Overland.Flow + Inlet.feature_Pipe + Inlet.feature_Other + Outlet.feature_Overland.Flow + Outlet.feature_Pipe + Outlet.feature_Riser.Pipe + Outlet.feature_Other + Engineered.soil.depth + s(Engineered.soil._..sand) + Engineered.soil._..silt + Engineered.soil._..clay + Engineered.soil._..compost + s(Engineered.soil._..other)

# Set the number of folds
k <- 6

# Create k equally sized folds
folds <- createFolds(y, k = k, list = TRUE, returnTrain = TRUE)

# Initialize a vector to store RMSE for each fold
rmse_values <- numeric(length(folds))

# Loop through each fold
for (i in seq_along(folds)) {
  # Split the data into training and testing sets
  test_indices <- folds[[i]]
  train_data <- df_res_removed[-test_indices, ]
  test_data <- df_res_removed[test_indices, ]
  
  # Fit the GAM model on the training data
  gam_model <- gam(formula, data = train_data)
  
  # Predict on the test data
  predictions <- predict(gam_model, newdata = test_data)
  
  # Calculate RMSE for this fold
  rmse_values[i] <- sqrt(mean((test_data$y - predictions)^2))
}

# Calculate the average RMSE across all folds
mean_rmse <- mean(rmse_values)
print(mean_rmse)
```

Cross Validation
```{r}
# Define response variable
y <- df_res_removed$Average.infiltration.rate

# Define training control with 5-fold cross-validation
train_control <- trainControl(method = "cv", number = 5)

# Train the GAM model using the train function
set.seed(123)  # For reproducibility
gam_model_caret <- train(
  y ~ s(Area) + s(Contributing.area) + Drainage.Area.Ratio+ s(Age) + s(Ponding.depth)+ Inlet.condition + Outlet.condition + Underdrain + X..Residential + s(X..Commercial) + X..Industrial + X..Institutional + s(X..Vegetation_shrub) + X..Vegetation_tree + X..Vegetation_grass + s(X..Vegetation_prairie)+ X..Vegetation_bare + s(X..Vegetation_other) +  Underlying.soil_..sand + Underlying.soil_..silt + Underlying.soil_..clay + Underlying.soil_..urban.disturbed + Underlying.soil_..other.undefined + s(Underlying.soil_..organic.matter) + Evidence.of.poor.infiltration + X..of.surface.with.poor.infiltration + Vegetation.condition + Inlet.feature_Curb.Cut + Inlet.feature_Overland.Flow + Inlet.feature_Pipe + Inlet.feature_Other + Outlet.feature_Overland.Flow + Outlet.feature_Pipe + Outlet.feature_Riser.Pipe + Outlet.feature_Other + Engineered.soil.depth + s(Engineered.soil._..sand) + Engineered.soil._..silt + Engineered.soil._..clay + Engineered.soil._..compost + s(Engineered.soil._..other), data = df_res_removed,
  method = "gam",
  trControl = train_control
)

# Print the results
print(gam_model_caret)
```

```{r}
# Example code

# Area+Contributing.area+Drainage.Area.Ratio+Age+Ponding.depth+Inlet.condition+Outlet.condition+Underdrain+X..Residential+X..Commercial+X..Industrial+X..Institutional+X..Vegetation_shrub+X..Vegetation_tree+X..Vegetation_grass+X..Vegetation_prairie+X..Vegetation_bare+X..Vegetation_other+Underlying.soil_..sand+Underlying.soil_..silt+Underlying.soil_..clay+Underlying.soil_..urban.disturbed+Underlying.soil_..other.undefined+Underlying.soil_..organic.matter+Evidence.of.poor.infiltration+X..of.surface.with.poor.infiltration+Vegetation.condition+Inlet.feature_Curb.Cut+Inlet.feature_Other+Inlet.feature_Overland.Flow+Inlet.feature_Pipe+Outlet.feature_Other+Outlet.feature_Overland.Flow+Outlet.feature_Pipe+Outlet.feature_Riser.Pipe+Engineered.soil.depth+Engineered.soil._..sand+Engineered.soil._..silt+Engineered.soil._..clay+Engineered.soil._..compost+Engineered.soil._..other

# Updated: Drainage.Area.Ratio + s(Age, bs = "re") + Ponding.depth + Underdrain + X..Residential  + s(X..Vegetation_shrub) + X..Vegetation_tree + s(X..Vegetation_grass) + s(X..Vegetation_prairie) + X..Vegetation_other + X..of.surface.with.poor.infiltration + Vegetation.condition + Engineered.soil.depth + s(Engineered.soil._..sand) + Engineered.soil._..compost + s(Soil_PC_1, bs = "re") + s(Soil_PC_2, bs = "re")

# Split into random training (75%) and validation (25%) sets
set.seed(123)
#random_indices <- sample(1:nrow(df_res_removed), 75)
#train_data <- df_res_removed[random_indices, ]
#validation_data <- df_res_removed[-random_indices, ]

# Fit GAM on training data
gam_model <- gam(Average.infiltration.rate ~ te(Drainage.Area.Ratio, X..Residential, bs = "re")  + s(X..Vegetation_shrub, X..Vegetation_tree) + s(X..Vegetation_grass, X..Vegetation_prairie) + X..of.surface.with.poor.infiltration + s(Engineered.soil._..sand, Engineered.soil._..compost) + s(Soil_PC_1, Soil_PC_2), method = "REML",  data = df_res_removed)

gam.check(gam_model)

summary(gam_model)

plot(gam_model, all.terms = TRUE, residuals=TRUE, pch=1, cex=1, shade=TRUE, shade.col="lightblue")

# Predict on validation data
#test_predictions <- predict(gam_model, newdata = validation_data)

# Predict on training data
#train_predictions <- predict(gam_model, newdata = train_data)

# Calculate performance metrics
#actual_values <- validation_data$Average.infiltration.rate

# Make plots square (better to visualize predictions when x and y axes are the same)
#par(pty="s")
# Plot actual vs predicted values for validation dataset
#plot(actual_values, test_predictions, xlab = "Actual", ylab = "Predicted", main = "Predicted vs. Actual Infiltration Rate", pch = 16, col = "dodgerblue2", bg="transparent", cex.axis=1)
#grid()
# Add the line y = x
#abline(a = 0, b = 1, col = "black", lwd = 2)

# Example: RMSE for training data
#train_rmse <- rmse(train_data$Average.infiltration.rate, train_predictions)
#print(paste("Training RMSE:", round(train_rmse, 4)))

# Example: R-squared for training data
#train_r2 <- cor(train_data$Average.infiltration.rate, train_predictions)^2
#print(paste("Training R-squared:", round(train_r2, 4)))

# Calculate performance metrics for testing data
#test_rmse <- rmse(validation_data$Average.infiltration.rate, test_predictions)
#print(paste("Testing RMSE:", round(test_rmse, 4)))

#test_r2 <- cor(validation_data$Average.infiltration.rate, test_predictions)^2
#print(paste("Testing R-squared:", round(test_r2, 4)))
```

Using gamclass for cross validation
```{r}
# Define response variable
y <- df_res_removed$Average.infiltration.rate

# Define the GAM model formula
formula <- Average.infiltration.rate ~ s(Area) + s(Contributing.area) + Drainage.Area.Ratio + s(Age) + s(Ponding.depth)+ Inlet.condition + Outlet.condition + Underdrain + X..Residential + s(X..Commercial, k=5) + X..Industrial + X..Institutional + s(X..Vegetation_shrub) + X..Vegetation_tree + X..Vegetation_grass + s(X..Vegetation_prairie)+ X..Vegetation_bare + s(X..Vegetation_other) +  Underlying.soil_..sand + Underlying.soil_..silt + Underlying.soil_..clay + Underlying.soil_..urban.disturbed + Underlying.soil_..other.undefined + s(Underlying.soil_..organic.matter) + Evidence.of.poor.infiltration + X..of.surface.with.poor.infiltration + Vegetation.condition + Inlet.feature_Curb.Cut + Inlet.feature_Overland.Flow + Inlet.feature_Pipe + Inlet.feature_Other + Outlet.feature_Overland.Flow + Outlet.feature_Pipe + Outlet.feature_Riser.Pipe + Outlet.feature_Other + Engineered.soil.depth + s(Engineered.soil._..sand) + Engineered.soil._..silt + Engineered.soil._..clay + Engineered.soil._..compost + s(Engineered.soil._..other)

# Perform cross-validation
cv_results <- CVgam(formula, data = df_res_removed, nfold = 5, debug.level = 0, method = "GCV.Cp",
              printit = TRUE, cvparts = NULL, gamma = 1, seed = 123)

# Print the cross-validation results
print(cv_results)

```
Investigate CVgam results
```{r}
# Assuming cv_results contains the output from CVgam
fitted_values <- cv_results$fitted
residuals <- cv_results$resid
actual_values <- df_res_removed$Average.infiltration.rate  # Ensure this matches your data

plot(actual_values, fitted_values)

```


Plot GAM Results
```{r}
df$predicted_y <- predict(gam_model, newdata = df_res_removed)
# Make plots square (better to visualize predictions when x and y axes are the same)
par(pty="s")
plot(df_res_removed$Average.infiltration.rate, df$predicted_y, main="",  xlab="Measured Infiltration Rate", ylab="Predicted Infiltration Rate", pch = 16, col = "salmon", bg="transparent", cex.axis=1)
abline(a = 0, b = 1, col = "black")
grid()
```

Best with REML, select=TRUE
s(Area) + s(Contributing.area) + s(Drainage.Area.Ratio) + s(Age) + s(Ponding.depth) + Inlet.condition + Outlet.condition + Underdrain + X..Residential + s(X..Commercial) + X..Industrial + s(X..Institutional) + s(X..Vegetation_shrub) + s(X..Vegetation_tree) + s(X..Vegetation_grass) + s(X..Vegetation_prairie) + X..Vegetation_bare + s(X..Vegetation_other) + s(Underlying.soil_..sand) + s(Underlying.soil_..silt) + s(Underlying.soil_..clay) + s(Underlying.soil_..urban.disturbed) + Underlying.soil_..other.undefined + s(Underlying.soil_..organic.matter) + Evidence.of.poor.infiltration + X..of.surface.with.poor.infiltration + Vegetation.condition + Inlet.feature_Curb.Cut + Inlet.feature_Overland.Flow + Inlet.feature_Pipe + Inlet.feature_Other + Outlet.feature_Overland.Flow + Outlet.feature_Pipe + Outlet.feature_Riser.Pipe + Outlet.feature_Other + s(Engineered.soil.depth) + s(Engineered.soil._..sand) + Engineered.soil._..silt + s(Engineered.soil._..clay) + Engineered.soil._..compost + s(Engineered.soil._..other), method="REML", select=TRUE,data = df_res_removed)

Best overall
s(Area) + s(Contributing.area) + Drainage.Area.Ratio+ s(Age) + s(Ponding.depth)+ Inlet.condition + Outlet.condition + Underdrain + X..Residential + s(X..Commercial) + X..Industrial + X..Institutional + s(X..Vegetation_shrub) + X..Vegetation_tree + X..Vegetation_grass + s(X..Vegetation_prairie)+ X..Vegetation_bare + s(X..Vegetation_other) +  Underlying.soil_..sand + Underlying.soil_..silt + Underlying.soil_..clay + Underlying.soil_..urban.disturbed + Underlying.soil_..other.undefined + s(Underlying.soil_..organic.matter) + Evidence.of.poor.infiltration + X..of.surface.with.poor.infiltration + Vegetation.condition + Inlet.feature_Curb.Cut + Inlet.feature_Overland.Flow + Inlet.feature_Pipe + Inlet.feature_Other + Outlet.feature_Overland.Flow + Outlet.feature_Pipe + Outlet.feature_Riser.Pipe + Outlet.feature_Other + Engineered.soil.depth + s(Engineered.soil._..sand) + Engineered.soil._..silt + Engineered.soil._..clay + Engineered.soil._..compost + s(Engineered.soil._..other), data = df_res_removed)



